### 2021-09-22 src/manuscript_figures.Rmd
update figure legends: replace latent factors with multi-omic fingerprints

---
title: "Manuscript Figures"
output:
  word_document: default
  html_document: default
params:
  settings_file: ''
  figures_folder: '/data/bimsbstatic/public/akalin/buyar/manuscript_figures_arcas/test'
  workdir: '.'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE, message = FALSE, 
                      fig.width = 9, fig.height = 11)
knitr::opts_knit$set(root.dir = params$workdir)
settings <- yaml::read_yaml(params$settings_file)
message(settings$GDCdataDir)
FigureFolder <- params$figures_folder
source(settings$utility_script)
library(data.table)
library(ggplot2)
library(pbapply)
library(ggrepel)
library(ggpubr)
library(cowplot)
library(ggridges)
library(ggcorrplot)
library(ComplexHeatmap)
library(pheatmap)
library(survminer)
library(survival)
library(caret)
library(treeheatr)
library(partykit)

theme_set(ggpubr::theme_pubclean())
geneSetScoresFolder <- '/data/local/buyar/arcas/subtyping_paper/data/geneset_scores/singscore'

inputDir <- settings$pipeline_output
assay_dir <- settings$assay_output$folder
surv <- do.call(rbind, readRDS(settings$surv))
projectDict <- sapply(yaml::read_yaml(settings$projectDict), 
                      function(x) gsub(' ', '', unlist(strsplit(x, ','))))
tcga_subtypes <- process_tcga_subtypes()

ens2hgnc <- readRDS(settings$ens2hgnc)
cpg2hgnc <- readRDS(settings$cpg2hgnc)

TOOLS <- c('maui_factors', 'mofa_factors', 'mcia_factors', 'pca_factors')

Figures <- list() # accummulate all figures in this
SuppFigures <- list() # accumulate all supplementary figures in this
Legends <- list() # figure legend text
SuppLegends <- list() # supplementary figure legend text

fig_count <- 0
supFig_count <- 0
```

```{r import_lfs}
LFs <- import_LFs(TOOLS, inputDir)
```

```{r import_feature_weights}
FWs <- import_feature_weights(TOOLS, inputDir)
```

```{r}
get_top_by_label <- function(dt, label.type, topN = 5) {
  do.call(rbind, lapply(split(dt, dt[,get(label.type)]), function(x) {head(x, topN)}))
}

map_ids <- function(markers, cpg2hgnc, ens2hgnc) {
  dt <- data.table::data.table('id' = gsub("^.+?\\.", "", markers), 
                               'type' = sub("(^.+?)\\..+$", "\\1", markers), 
                               'variable' = markers)
  do.call(rbind, lapply(names(split(dt, dt$type)), function(x) {
    message(x)
    if(x == 'meth') {
      dt.map <- cpg2hgnc[cpg_site %in% dt[type == x]$id]
      colnames(dt.map) <- c('id', 'hgnc')
      return(merge(dt[type == x], dt.map,  by = 'id'))
    } else if (x == 'cnv') {
      dt.map <- dt[type == x]
      dt.map$hgnc <- gsub("^.+?\\.", "", dt.map$variable)
      return(dt.map)
    } else if (x == 'gex' || x == 'mut') {
      dt.map <- ens2hgnc[ref_gene_id %in% dt[type == x]$id]
      colnames(dt.map) <- c('id', 'hgnc')
      return(merge(dt[type == x], dt.map,  by = 'id'))
    }
  }))
}
```

```{r pancancer_prediction}
# In this section, we make figures about pan-cancer classification 
# and some benchmarks of cohort-based unsupervised clustering across cancers

# 1. pancancer classification
# Find out if cancer types are predictable using latent factors 
res <- sapply(simplify = F, names(LFs), function(tool) {
  message(date(), " => ", tool)
  M <- LFs[[tool]]$pancancer.cnv_gex_meth_mut
  ct <- gsub("-", "_", surv[match(rownames(M), bcr_patient_barcode)]$project)
  res <- predict_labels_glmnet_multiclass(M, ct, partition = 0.6, nodes = 20)
})

pancancer_stats <- sapply(res, function(x) {
  caret::multiClassSummary(x, lev = levels(x$obs))
})

dt <- data.table(melt(pancancer_stats))[Var1 %in% c('Mean_Balanced_Accuracy')]
dt$Var2 <- toupper(gsub("_factors", "", dt$Var2))
p1 <- ggplot(dt, aes(x = Var2, y = value)) + 
  geom_bar(stat = 'identity', aes(fill = Var2)) + 
  labs(y = 'Mean Balanced Accuracy') + 
  theme(legend.position = 'none', axis.title.x = element_blank(), 
        axis.text = element_text(size = 14), axis.title = element_text(size = 16)) + 
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1))

# performance visualisation - misclassified examples 
m <- as.matrix(table(res$maui_factors$pred, res$maui_factors$obs))
# compute percentage of correct predictions
hm <- apply(m, 1, function(x) round(x/sum(x) * 100))
hm_labels <- hm
hm_labels[hm_labels < 2] <- '' #don't display very low values
p2 <- pheatmap::pheatmap(hm, cluster_rows = F, cluster_cols = F, 
          silent = T, cellwidth = 10, cellheight = 10, color = colorRampPalette(c('white', 'red'))(10), 
          display_numbers = hm_labels, fontsize = 8)

# plot cancer specific factors
M <- LFs$maui_factors$pancancer.cnv_gex_meth_mut
ct <- surv[match(rownames(M), bcr_patient_barcode)]$project
dt <- get_factor_specific_variables(M, ct)[padj < 0.05][order(padj)]
top_vars <- get_top_by_label(dt, 'ref_cl', topN = 5)

p3 <- plot_tsne(M[,unique(top_vars$variable)], ct, show.labels = T) + 
  theme(legend.position = 'none')

p4 <- plot_label_specific_vars(M, ct)

# data type contribution to top LF per factor
p5 <- plot_datatype_contribution_by_factor(LFs, FWs, 'maui_factors', 
                                           'pancancer', 'cnv_gex_meth_mut', factors = ct, 
                                           top_lfs = 1, top_features = 100) 
p5 <- p5 + labs(y = 'Relative Contribution\nto top latent factor') + 
  theme(legend.position = 'right', legend.direction = 'vertical', 
        legend.background = element_blank(), 
        legend.title = element_blank(), legend.box.just = 'left')

# 2. Cohort-based unsupservised clustering benchmark
# clustering metrics of various cancers 
# Use top most-variable factors (derived from multi-omics) to cluster samples using cancer-specific LFs 
multi_omics_comb <- 'cnv_gex_meth_mut'
experiments <- grep(paste0('^TCGA-.*.', multi_omics_comb), names(LFs$pca_factors), value = T)
cl <- parallel::makeCluster(20)
parallel::clusterExport(cl, varlist = c('LFs', 'experiments', 'settings'))
cluster.stats <- do.call(rbind, sapply(simplify = F, names(LFs), function(tool) {
  message("Clustering top ",tool)
  stats <- do.call(rbind, pbapply::pbsapply(cl = cl, simplify = F, 
                                            intersect(names(LFs[[tool]]), experiments), 
                    function(experiment) {
    message(experiment)
    source(settings$utility_script)
    library(data.table)
    M <- LFs[[tool]][[experiment]]
    if(tool == 'maui_factors') {
      # to avoid highly correlated maui LFs to dominate clusters
      M <- scale(t(remove_redundant_variables(t(M))))
    } 
    # use top 5 most-variable factors for clustering
    M.sub <- t(subset_features_by_variance(t(M), topN = 5))
    val <- c('internal') 
    if(ncol(M.sub) > 2) {val <- c('internal', 'stability')}
    cl_res <- clValid::clValid(obj = M.sub,
                          nClust = 3:8,
                          clMethods = 'kmeans',
                          validation = val,
                          maxitems = 10000,
                          nstart = 1000, iter.max = 30)
    stats <- data.table::as.data.table(cl_res@measures)[V1 %in% c('Silhouette', 'APN')]
    colnames(stats) <- c('metric', 'k', 'clustering_method', 'score')
    stats$cancer_type <- strsplit(experiment, "\\.")[[1]][1]
    stats$datatypes <- strsplit(experiment, "\\.")[[1]][2]
    return(stats)
  }))
  stats$method <- tool
  return(stats)
}))
parallel::stopCluster(cl)

# plot silhouette/stability score comparison
dt1 <- cluster.stats[metric == 'Silhouette'][datatypes == multi_omics_comb,
                                             .SD[which.max(score)],
                                             by = c('cancer_type', 'method', 'k')]
dt1$method <- toupper(gsub("_factors", "", dt1$method))
comps <- lapply(setdiff(unique(dt1$method), 'MAUI'), function(x) c('MAUI', x))

p6 <- ggboxplot(dt1, 
          x = 'method', y = 'score', add = 'jitter', color = 'method') + 
  stat_compare_means(method = 'wilcox.test', paired = TRUE, 
                     comparisons = comps) + 
  labs(title = 'Silhouette score', x = '') + 
  theme(legend.position = 'none')

dt2 <- cluster.stats[metric == 'APN'][datatypes == multi_omics_comb,
                                             .SD[which.min(score)],
                                             by = c('cancer_type', 'method', 'k')]
dt2$method <- toupper(gsub("_factors", "", dt2$method))
comps <- lapply(setdiff(unique(dt2$method), 'MAUI'), function(x) c('MAUI', x))
p7 <- ggboxplot(dt2, 
          x = 'method', y = 'score', add = 'jitter', color = 'method') + 
  stat_compare_means(method = 'wilcox.test', paired = TRUE,  
                     comparisons = comps) + 
  labs(title = 'Cluster Stability Score (APN)', x = '') + 
  theme(legend.position = 'none')

# combine plots for main figure 
Fig <- plot_grid(plot_grid(p3, p5, nrow = 1, labels = c('A', 'B'),
                           rel_widths = c(2,1.5)), 
                  plot_grid(p6, p7, 
                            nrow = 1, labels = c('C','D')), 
                  nrow = 2,
                  rel_heights = c(2,1))

fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.benchmarks.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name),
              plot = Fig, units = 'in', width = 12, height = 12)
Legends[[fig_count]] <- " A) tSNE plot of MAUI multi-omics fingerprints (using top 5 fingerprint enriched per cancer type) learned from pan-cancer dataset  B) Percentage of top 100 features by omics types contributing to top predictive MAUI fingerprint per cancer type. C, D) Comparison of unsupervised clustering performances based on cluster homogeneity (Silhouette) in C and cluster stability (Average Proportion of Non-Overlap (APN) in D. Each point represents a score for an individual clustering experiment in a total of 132 k-means clustering experiments (21 cancer types x for 6 different values of k from 3 to 8) for each tool."

# supplementary figure
supFig <- cowplot::plot_grid(cowplot::plot_grid(p1, p2$gtable, nrow = 1, 
                                                labels = c('A', 'B'), rel_widths = c(1.2, 2)), 
                          p4, nrow = 2, labels = c('', 'C'), rel_heights = c(1.2, 2))
# print p3 to supplements 
supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- supFig
fig.name <- paste0('SuppFigure', supFig_count, '.benchmarks.supp.pdf')
names(SuppFigures)[supFig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = supFig, units = 'in', width = 8.27, height = 11.69)
SuppLegends[[supFig_count]] <- "A) Predictive power of latent factors in distinguishing cancer types comparing MAUI with PCA/MOFA/MCIA. Mean Balanced Accuracy computed on the test dataset (30% of held-out data) trained on 70% of the dataset using an Elastic Net multi-class model. B) Confusion matrix of predicted versus true labels of samples' cancer types. The color of each cell represents the percentage of samples from the cancer type in the rows that are classified into the cancer types on the column. Cells that deviate from the red color represent relatively lower
level of classification accuracy. C) Top MAUI latent factors (multi-omic fingerprints) detected per cancer type. Size/color/opacity of the points represent the average neural activation value of the MAUI fingerprint for the corresponding type of tumor samples."
```

```{r clinical_survival_factors}
# here we compute LFs associated to clinical factors (e.g. tumor stage, gender etc) and 
# survival outcomes (PFI)

## 1. clinical factor associated MAUI LFs
# Find maui LFs associated to clinical factors and plot them
experiments <- grep('^TCGA-.*.cnv_gex_meth_mut', names(LFs$maui_factors), value = T)
# find clinical-factor associated LFs 
cl <- parallel::makeCluster(10)
parallel::clusterExport(cl, varlist = c('experiments', 'LFs', 'settings'))
clin.stats <- do.call(rbind, 
        pbapply::pbsapply(cl = cl, simplify = F, 
               experiments, function(experiment) {
               require(data.table)
               source(settings$utility_script)
               M <- LFs$maui_factors[[experiment]]
               clin <- process_clinical_covariates(rownames(M))
               dt <- do.call(rbind, 
                    pbapply::pbsapply(simplify = F, 
                                      colnames(clin), function(x) {
                      if(!is.numeric(clin[,x])) {
                        factors <- clin[,x]
                        m <- M[!is.na(factors),]
                        factors <- factors[!is.na(factors)]
                        dt <- get_factor_specific_variables(m, factors) 
                        dt$label <- x
                        return(dt)
                      }
                      return(NULL)
                      }))
               dt$project <- unlist(strsplit(experiment, "\\."))[[1]]
               return(dt)
             }))
parallel::stopCluster(cl)

# find top factors per project per label and plot
dt <- clin.stats[padj < 0.05, .SD[which.min(padj)],by = c('project', 'label')]

require(ggrepel)
dt$label <- stringr::str_wrap(gsub('_', ' ', dt$label), width = 8)
p1 <- ggplot(dt, aes(x = project, y = label)) + 
  geom_point(aes(color = -log10(padj), size = -log10(padj))) + 
  geom_text_repel(aes(label = variable)) + 
  coord_flip() +
  scale_color_gradient(low = 'blue', high = 'red') + 
  labs(x = '', y = 'clinical factor')

# 2. survival-predictive factors
#I look for survival-predictive maui LFs, show how much c-index improves on top of 
#clinical factors (tumor stage/age/gender) and plot top survival-predictive factor per cancer.

# factors to use in the model
clin_factors <- c('age_at_initial_pathologic_diagnosis', 'gender', 'ajcc_pathologic_tumor_stage')
event <- 'PFI'
time <- 'PFI.time'

experiments <- grep('^TCGA-.*.cnv_gex_meth_mut', names(LFs$maui_factors), value = T)
cl <- parallel::makeCluster(10)
parallel::clusterExport(cl, varlist = c('LFs', 'clin_factors', 'event', 'time',
                                        'surv', 'experiments', 'settings'))
surv.stats <- pbapply::pbsapply(cl = cl, simplify = F, experiments, function(experiment) { 
  require(data.table)
  require(survminer)
  require(survival)
  source(settings$utility_script)
  message(experiment)
  M <- LFs$maui_factors[[experiment]]
  # get clinical factors for the patients in matrix
  clin <- process_clinical_covariates(rownames(M))
  clin <- subset(clin, select = intersect(colnames(clin), clin_factors))
  # first filter variables by importance 
  fit <- fit_RSF(surv, M, 500, time, event)
  vars <- names(head(sort(fit$importance, decreasing = T), 5))

  # compute adjusted survival curve for thb top factor
  v <- vars[1] 
  dat <- merge(clin, M[,v,drop = F], by = 'row.names')
  rownames(dat) <- dat$Row.names
  dat$Row.names <- NULL
  dat <- get_surv_df(surv, dat, event, time)

  # split variable into high/low to compute survival adjusted curves
  cp <- find_survival_cutpoint(dat[,v,drop=F], surv, event_col = event, duration_col = time)
  dat[,v] <- ifelse(dat[,v] > cp, paste(v, '>', round(cp, 1)), paste(v, '<=', round(cp, 1)))
  cox2 <- coxph(Surv(time,status) ~ ., 
                data=dat, x = TRUE)
  p <- survminer::ggadjustedcurves(fit = cox2, data = dat, variable = v, method = 'average')

  # compute Cindex with all survival-predictive LFs and without LFs 
  rsf_stats <- fit_RSF_clin(get_surv_df(surv, clin, event, time), 
                            M[,vars], ntree = 1000)

  return(list('p' = p, 
              'Cindex_base' = rsf_stats$C_wo, 
              'Cindex_with_LFs' = rsf_stats$C_with))
})
parallel::stopCluster(cl)

# make a plot to show improvement in C index with LFs on top of clinical factors
ci <- do.call(rbind, lapply(names(surv.stats), function(x) {
  if(is.null(surv.stats[[x]])) {return(NULL)}
  df <- data.frame('base' = surv.stats[[x]][['Cindex_base']], 
                   'with_lf' = surv.stats[[x]][['Cindex_with_LFs']], 
                   'project' = unlist(strsplit(x, '\\.'))[[1]])
  return(df)
}))

dt <- melt(ci, id.vars = 'project')
dt$variable <- ifelse(dt$variable == 'base', 'Age+gender+stage', 'Age+gender+stage + MAUI LFs')

p2 <- ggplot(ci, aes(x = reorder(project, with_lf))) + 
  geom_segment(aes(x = reorder(project, with_lf), xend = project, 
                   y = base, yend = with_lf)) + 
  geom_hline(yintercept = c(0.5, 0.6, 0.7), linetype = 2) + 
  geom_point(data = dt, aes(x = project, y = value, color = variable), size = 2) +
  coord_flip() + 
  labs(x = '', y = '') + 
  theme(legend.title = element_blank(), legend.position = 'bottom', 
        legend.background = element_blank(), 
        legend.margin = margin(), 
        legend.direction = 'vertical', legend.box.background = element_blank()) +
  scale_color_brewer(type = 'qual', palette = 6)

Fig <- cowplot::plot_grid(cowplot::plot_grid(p1, p2, nrow = 2, labels = c('A', 'B')))

fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.clinical_factors.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(filename = file.path(FigureFolder, fig.name), 
       plot = Fig, width = 8.27, height = 11.69, units = 'in')
Legends[[fig_count]] <- "A) Top significant (padj < 0.05) clinical factor-associated multi-omic fingerprints per cancer type. B) Harrell’s C-index for progression-free survival prediction accuracy computed using only clinical factors (age + gender + tumor stage) or using clinical factors along with survival-predictive MAUI fingerprints"

# combine adjusted survival curve plots into a single plot 
plots <- sapply(simplify = F, names(surv.stats), function(x) {   
  p <- surv.stats[[x]][['p']] 
  if(!is.null(p)) { 
    p + theme(axis.text.x = element_text(size = 8), 
              axis.text.y = element_text(size = 8), 
              axis.title.x = element_blank(), 
              axis.title.y = element_blank(), 
              legend.title = element_blank(),
              legend.position = c(0.5, 0.9),
              legend.text = element_text(size = 8), 
              legend.background = element_blank(), 
              legend.direction = 'vertical') 
    }
})
plots <- plots[!sapply(plots, is.null)]
Fig <- cowplot::plot_grid(plotlist = plots, nrow = 6, 
                   labels = sapply(strsplit(names(plots), '\\.'), function(x) x[1]), 
                   label_size = 8)

# print p3 to supplements 
supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- Fig
fig.name <- paste0('SuppFigure', supFig_count, '.adjusted_survival_curves.pdf')
names(SuppFigures)[supFig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = Fig, units = 'in', width = 8.27, height = 11.69)
SuppLegends[[supFig_count]] <- "Kaplan-Meier survival curves adjusted for clinical factors (tumor stage + age + gender) stratified by the neural activation values of top survival-predictive multi-omic fingerprints per cancer."
```

```{r tumor_purity} 
experiments <- grep('TCGA.*cnv_gex_meth_mut', names(LFs$maui_factors), value = T)
# use maui LFs to predict TCGA IHC/CPE values that represent tumor purity levels 
cl <- parallel::makeCluster(20)
parallel::clusterExport(cl, varlist = c('LFs', 'experiments', 'settings'))
results <- pbapply::pbsapply(cl = cl, simplify = F, experiments, function(experiment) {
    message(experiment)
    source(settings$utility_script)
    library(data.table)
    M <- LFs$maui_factors[[experiment]]
    tp <- get_tumor_purity_estimates(rownames(M))
    if(nrow(tp) < 50) {
      warning("Couldn't find at least 100 samples with tumor purity 
              scores in experiment ", experiment, " => ", nrow(tp))
      return(NULL)
    }
    # predict IHC
    y1 <- tp[,mean(IHC), by = bcr_patient_barcode][match(rownames(M), bcr_patient_barcode)]$V1
    res1 <- predict_numerical_glmnet(M, y1)
    # predict CPE (consensus estimates)
    y2 <- tp[,mean(CPE), by = bcr_patient_barcode][match(rownames(M), bcr_patient_barcode)]$V1
    res2 <- predict_numerical_glmnet(M, y2)
    return(list('IHC' = res1, 'CPE' = res2))
})
parallel::stopCluster(cl)

# summarize stats 
stats <- do.call(rbind, lapply(names(results), function(x) {
  message(x)
  if(is.null(results[[x]])) {
    return(NULL)
  }
  dt <- do.call(rbind, lapply(names(results[[x]]), function(y) {
    message(y)
    dt <- evaluate_regression_model(results[[x]][[y]][['preds']]$y, 
                              results[[x]][[y]][['preds']]$y_hat)
    dt$purity_method <- y
    return(dt)
  }))
  dt$experiment <- x
  return(dt)
}))

stats$project <- unlist(lapply(strsplit(stats$experiment, "\\."), function(x) x[[1]]))

# plot correlation scores for all cohorts 
p1 <- ggplot(stats[purity_method == 'IHC'][grep('TCGA', project)], 
       aes(x = reorder(project, COR), y = COR, fill = COR),) + 
  geom_bar(stat = 'identity') +
  coord_flip() +
  theme(axis.title.y = element_blank(), legend.position = 'none') + 
  labs(y = "MAUI vs IHC") + ylim(0, 1)

# compute other tools' correlations with IHC
tp <- get_tumor_purity_estimates()
absolute <- tp[!is.na(ABSOLUTE), mean(cor(ABSOLUTE, IHC, use = 'complete.obs')),by = project]
estimate <- tp[!is.na(ESTIMATE), mean(cor(ESTIMATE, IHC, use = 'complete.obs')),by = project]
lump <- tp[!is.na(LUMP), mean(cor(LUMP, IHC, use = 'complete.obs')),by = project]

dt <- stats[purity_method == 'IHC']
colnames(dt)[3] <- 'MAUI'
dt$ABSOLUTE <- absolute[match(dt$project, project)]$V1
dt$LUMP <- lump[match(dt$project, project)]$V1
dt$ESTIMATE <- estimate[match(dt$project, project)]$V1

mdt <- melt(dt, measure.vars = c('MAUI', 'ABSOLUTE', 'ESTIMATE', 'LUMP'), id.vars = 'project')
p2 <- ggboxplot(mdt, x = 'variable', y = 'value', add = 'jitter', color = 'variable') +
  theme(legend.position = 'none', axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(y = 'Average Correlation with IHC')

p3 <- ggplot(stats[purity_method == 'CPE'][grep('TCGA', project)], 
       aes(x = reorder(project, COR), y = COR, fill = COR)) + 
  geom_bar(stat = 'identity') +
  coord_flip() +
  theme(axis.title.y = element_blank(),  legend.position = 'none') + 
  labs(y = "MAUI vs CPE") + ylim(0, 1)

# correlation plots for top cancer types
selected <- stats[purity_method == 'IHC'][grep('TCGA', experiment)][order(COR, decreasing = T)][1:2]$experiment
plots <- lapply(selected, function(x) {
  message(x)
  p <- cowplot::plot_grid(plotlist = lapply(names(results[[x]]), function(y) {
    message(y)
    ggpubr::ggscatter(results[[x]][[y]][['preds']], 
                           x = 'y_hat', y = 'y', 
                           add = 'reg.line', cor.coef = T, cor.method = 'pearson') + 
      labs(x = paste('MAUI prediction of', y), y = y, title = gsub("\\..+$", "", x))
  }))
  return(p)
})

p4 <- cowplot::plot_grid(plotlist = plots, ncol = 1)

# supplementary figure
supFig <- cowplot::plot_grid(cowplot::plot_grid(p1, p2, p3, labels = c('A', 'B', 'C'), nrow = 1), 
                             p4, nrow = 2, labels = c('', 'D'))

# print p3 to supplements 
supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- supFig
fig.name <- paste0('SuppFigure', supFig_count, '.tumor_purity.supp.pdf')
names(SuppFigures)[supFig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = supFig, units = 'in', width = 8.27, height = 11.69)
SuppLegends[[supFig_count]] <- "Predicting tumor purity estimates using MAUI latent factors. 
A) Correlation of the MAUI predictions of IHC versus the annotated IHC values (immunohistochemistry-based tumor purity estimation). 
B) Comparison of average correlations between computational estimates of IHC values by different tools. Each point represents the 
value obtained for each of the TCGA cohorts seen in panel A. ABSOLUTE tool lacks estimates for 6 out of 14 cancer types. 
C) Correlation of the MAUI predictions of CPE versus the annotated CPE levels (CPE stands for consensus purity estimated by computational methods: ABSOLUTE, ESTIMATE, and LUMP). 
D) Scatterplots of MAUI predictions versus the annotated IHC and CPE values for the 
top two best predicted cancer types with respect to annotated IHC values."
```

```{r msi_stats}
# Predicting MSI status using latent factors
# predict MSI status for pan-gastrointestinal tumors
# read MSI data
dataDir <- settings$GDCdataDir
MSI <- do.call(rbind, lapply(projectDict$`pan-gastrointestinal`, function(pr) {
  f <- file.path(dataDir, paste0(pr, '.MSI.tsv'))
  dt <- data.table::fread(f)[,c(2,4)]
  colnames(dt)[2] <- 'MSI'
  dt$project <- pr
  return(dt)
}))
# remove samples with undertermined MSI status
MSI <- MSI[MSI %in% c('MSI-H', 'MSI-L', 'MSS')]

dt <- do.call(rbind, lapply(names(LFs), function(tool) {
  M <- LFs[[tool]][['pan-gastrointestinal.cnv_gex_meth_mut']]
  y <- MSI[match(rownames(M), bcr_patient_barcode)]$MSI
  M <- M[!is.na(y),]
  y <- y[!is.na(y)]

  # find LFs predictive of MSI 
  dt <- get_factor_specific_variables(M, y)
  dt$tool <- tool
  return(dt)
}))

# compare top predictive factors across tools
dt$tool <- toupper(gsub("_factors", "", dt$tool))
p1 <- ggplot(dt[,.SD[which.min(padj)],by = c('ref_cl', 'tool')], 
       aes(x = ref_cl, y = -log10(padj))) + 
  geom_bar(stat = 'identity', aes(fill = tool), position = 'dodge') +
  geom_hline(yintercept = -log10(0.05)) + 
  coord_flip() + 
  theme(legend.title = element_blank(), axis.title.y = element_blank(), 
        plot.title = element_text(hjust = 0.5))

# focus on maui LFs to visualize MSI status-related LFs
M <- LFs$maui_factors[['pan-gastrointestinal.cnv_gex_meth_mut']]
y <- MSI[match(rownames(M), bcr_patient_barcode)]$MSI
M <- M[!is.na(y),]
y <- y[!is.na(y)]
top_vars <- unique(get_top_by_label(dt[tool=='MAUI'][padj < 0.05][order(padj)], 'ref_cl', 10)[['variable']])
df <- plot_tsne(M[,top_vars], y, returnData = T)
top_lf <- dt[order(padj)][ref_cl == 'MSI-H']$variable[1]
df <- merge(df, data.frame(M[,top_lf,drop =F]), by = 'row.names')
p2 <- ggplot(df, aes(x = tSNE1, y = tSNE2)) + 
    geom_point(aes(color = as.factor(group))) + 
    theme(legend.title = element_blank())
p3 <- ggplot(df, aes(x = tSNE1, y = tSNE2)) + 
    geom_point(aes_string(color = top_lf)) + 
    scale_color_gradient(low = 'gray', high = 'red')

p4 <- ggboxplot(df, x = 'group', y = top_lf, color = 'group', 
          add = 'jitter') + theme(legend.position = 'none', 
                                  axis.title.x = element_blank())

# plot a roc curve
rocs <- predict_cluster_labels_glmnet(M, y, returnROC = TRUE)
p5 <- plot_roc_list(rocs)

# top features that contribute to top factors relevant to MSI status prediction
f <- data.table::fread(file.path(inputDir, 'maui_factors', 
                               paste('pan-gastrointestinal', # project 
                                      'cnv_gex_meth_mut', #data type
                                      'maui_factors', # tool
                                      'feature_weights.csv', sep = '.')))
feat.M <- as.matrix(f[,-1])
rownames(feat.M) <- gsub('assay: ', '', f$V1)

# sort features by contribution to LF and plot by data type 
df <- data.frame(feat.M[,top_lf,drop = F])
df$data.type <- sub('^(.+?)\\..+$', '\\1', rownames(df))

# get top 100 features
df <- df[order(abs(df[,top_lf]), decreasing = T),]
freq <- do.call(rbind, lapply(0:29, function(i) {
  s <- (i*100+1)
  e <- (i+1)*100
  x <- data.frame(table(df[s:e,]$data.type))
  x$rank <- stringr::str_wrap(paste0("[",s,",",e,"]"),
                              width = 8)
  x$ind <- i
  return(x)
}))
p6 <- ggplot(freq[freq$ind < 5,], aes(x = reorder(rank, ind), y = Freq)) + 
  geom_bar(aes(fill = Var1), stat='identity') + 
  theme(legend.title = element_blank(), axis.text.x = element_text(angle = 45)) + 
  labs(x = paste0("top features contributing to ",top_lf))

Fig <- cowplot::plot_grid(p1, p5, p2, p3, p4, p6,
                          nrow = 3, labels = 'AUTO')
fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.MSI.status.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(filename = file.path(FigureFolder, fig.name), 
       plot = Fig, units = 'in', width = 12, height = 12)
Legends[[fig_count]] <- "Latent factors predictive of MSI status (high: MSI-H, stable: MSI-S, low: MSI-L) in pan-gastrointestinal cancers A) Comparison of top latent factor per MSI status for each tool in terms of adjusted p-values reflecting the differences between MSI subgroups. B) ROC curves for prediction accuracy of MSI status using MAUI latent factors (multi-omic fingerprints) C) tSNE plots of samples generated using top MSI-predictive fingerprints (5 LF per MSI class) colored by MSI status D) tSNE plots of samples colored by MAUI fingerprint LF25, which is most predictive of MSI-H samples. E) Distribution of MAUI fingerprint LF25 across MSI subgroups in the pan-gastrointestinal cancer samples. F) Relative abundance of omics feature types among the top 500 features contributing to MAUI fingerprint LF25."
```

```{r}
# Analysis of non-small-cell lung cancer types (LUSC vs LUAD)
# reference study: https://www.nature.com/articles/s41598-020-77284-8
M <- LFs$maui_factors$nsclc.cnv_gex_meth_mut
M <- scale(M) 
ct <- gsub("TCGA-", "", surv[match(rownames(M), bcr_patient_barcode)]$project)

# find type-specific LFs
subtype_lfs <- get_factor_specific_variables(M, ct)
top_vars <- get_top_by_label(subtype_lfs[padj < 0.05][order(padj)], 'ref_cl', 1)

# roc curves for predicting lung cancer types
rocs <- predict_cluster_labels_glmnet(M, ct, returnROC = TRUE)
aucs <- sapply(rocs, pROC::auc)
p1 <- pROC::ggroc(rocs, size = 1)
p1$data$name <- paste0(p1$data$name, " (AUC=", round(aucs[p1$data$name], 3),")")
p1 <- p1 + geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color="grey", linetype="dashed") +
  theme(legend.title = element_blank(), legend.direction = 'vertical')

# tsne plot 
df <- plot_tsne(M, ct, returnData = T)

# top lf for LUSC
top_lf_lusc <- top_vars[ref_cl == 'LUSC']$variable
top_lf_luad <- top_vars[ref_cl == 'LUAD']$variable
df <- cbind(df, 
            M[,top_lf_lusc,drop = F],
            M[,top_lf_luad,drop = F])
p <- ggplot(df, aes(x = tSNE1, y = tSNE2)) 
p2 <- p + geom_point(aes(color = group)) + 
  geom_text_repel(data = get_centroid(df[,1:2], df$group), 
            aes(x,y,label = text.label), size = 5)
p3 <- p + geom_point(aes_string(color = top_lf_lusc)) + 
  scale_color_gradient(low = 'gray', high = 'red') +
  geom_text_repel(data = get_centroid(df[,1:2], df$group), 
            aes(x,y,label = text.label), size = 5)
p4 <- p + geom_point(aes_string(color = top_lf_luad)) + 
  scale_color_gradient(low = 'gray', high = 'red') +
  geom_text_repel(data = get_centroid(df[,1:2], df$group), 
            aes(x,y,label = text.label), size = 5)

# detect subtype specific features that contribute most 
# to top LFs and make a decision tree plot  
## get top 10 markers per top LFs per group
top_vars <- get_top_by_label(subtype_lfs[padj < 0.05][order(padj)], 'ref_cl', 5)
W <- FWs$maui_factors$nsclc.cnv_gex_meth_mut
markers <- unlist(lapply(unique(top_vars$variable), function(x) {
  # get top features per lf and concatenate
  head(sort(abs(W[,x]), decreasing = T), 10)
}))

# import assay data and subset for markers 
A <- data.table::fread(file.path(settings$assay_output$folder, 
                                 'nsclc.cnv_gex_meth_mut.csv'))
A.sub <- A[V1 %in% names(markers)]
#convert to matrix
A.subM <- as.matrix(data.frame(A.sub[,-1],
                               row.names = A.sub$V1, 
                               check.names = F))
labels <- gsub("TCGA-", "", surv[match(colnames(A.subM), bcr_patient_barcode)]$project)
# use treeheatr and partykit package to compute/visualize decision tree

top_markers <- map_ids(names(markers), cpg2hgnc, ens2hgnc)
# get unique table 
top_markers <- top_markers[match(top_markers[,head(variable, 1), by = c('type', 'hgnc')]$V1, variable)]

df <- data.frame(t(round(A.subM[top_markers$variable,], 3)))
colnames(df) <- paste(top_markers$type, top_markers$hgnc, sep = ":")
df$subtype <- as.factor(labels)

set.seed(123)
inTrain <- caret::createDataPartition(df$subtype, p = 0.6)
df.train <- df[inTrain[[1]],]
df.test <- df[-inTrain[[1]],]
# further filter top features: build initial tree to get subset of most important features
x = partykit::ctree(subtype ~ ., data = df.train)
selected <- names(partykit::varimp(x))
df.train <- df.train[,c(selected, 'subtype')]
df.test <- df.test[,c(selected, 'subtype')]
p5 <- treeheatr::heat_tree(
  x = partykit::ctree(subtype ~ ., data = df.train),
  data_test = df.test,
  heat_rel_height = 0.4
)

# compute classification accuracy of using top 10 markers in a random forest model
rf_stats <- predict_cluster_labels(t(A.subM[top_markers$variable,]), labels, partition = 0.6)

Fig <-  cowplot::plot_grid(cowplot::plot_grid(p1, p2, p3, p4,
                         labels = 'AUTO', nrow = 2), 
                         p5, nrow = 2, labels = c('', 'E'))

fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.luad_vs_lusc.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(filename = file.path(FigureFolder, fig.name), 
       Fig, units = 'in', 
       width = 8.27, height = 11.69)

Legends[[fig_count]] <- "Classification of Non-Small-Cell Lung Adenocarcinoma samples using MAUI multi-omic fingerprints.
A) ROC curves with AUC values on testing data (40% of the dataset) using an Elastic Net model with 5-fold cross-validation. B) t-SNE plot of MAUI fingerprints colored by cancer types. C) t-SNE plot of MAUI fingerprints colored by 
the top most-predictive fingerprint for TCGA-LUSC samples. D) t-SNE plot of MAUI fingerprints colored by the top 
most predictive fingerprint for TCGA-LUAD samples. E) Decision tree visualisation of top MAUI fingerprint-associated omics markers detected that can be used to distinguish LUSC and LUAD subtypes."
```

```{r immunotherapy_response}
# immunotherapy response prediction analysis: 
# TODO: pass paths via args
projDir <- '/data/local/buyar/arcas/pancancer_multiomics_manuscript/results/immunotherapy_response/Mariathasan_et_al'
settings_imtrp <- yaml::read_yaml(file.path(projDir, 'settings.yaml'))
gex_all <- readRDS(file.path(projDir, 'mariathasan.gex.all_genes.normalised.RDS'))
human_gene_signatures <- readRDS(file.path(projDir, 'human_gene_signatures.RDS'))

LFs_imtrp <- import_LFs(c('maui_factors', 'mofa_factors', 'pca_factors'), settings_imtrp$pipeline_output)
FWs_imtrp <- import_feature_weights(c('maui_factors'), settings_imtrp$pipeline_output)
colData <- data.table::fread(settings_imtrp$clin)
colnames(colData)[1] <- 'bcr_patient_barcode'

dt <- do.call(rbind, lapply(names(LFs_imtrp), function(tool) {
  M <- LFs_imtrp[[tool]]$mariathasan.gex
  # find LFs predictive of response to therapy 
  dt <- get_factor_specific_variables(M, colData$binaryResponse)[order(padj)]
  dt$tool <- tool
  return(dt)
}))

dt$tool <- toupper(gsub("_factors", "", dt$tool))
dt$ref_cl <- ifelse(dt$ref_cl == 'CR/PR', 'Responders', 'Non-responders')

p1 <- ggplot(dt[,.SD[which.min(padj)],by = c('tool', 'ref_cl')], 
       aes(x = ref_cl, y = -log10(padj))) + 
  geom_bar(stat = 'identity', aes(fill = tool), position = 'dodge') +
  geom_hline(yintercept = -log10(0.05)) + 
  coord_flip() + 
  theme(legend.title = element_blank(), axis.title.y = element_blank(), 
        plot.title = element_text(hjust = 0.5), 
        legend.key.size = unit(0.2, units = 'in'))

# correlation of top LF with tumor mutation burden
# get samples with both response and tmb annotation
dt <- do.call(rbind, lapply(names(LFs_imtrp), function(tool) {
  M <- LFs_imtrp[[tool]]$mariathasan.gex
  dt <- data.table(cor(M, colData$`Neoantigen burden per MB`, use = 'complete.obs'),
                   keep.rownames = T)
  dt$tool <- tool
  return(dt)
}))
dt <- dt[,.SD[which.max(abs(V1))], by = c('tool')]
dt$tool <- toupper(gsub("_factors", "", dt$tool))
p2 <- ggplot(dt, aes(x = tool, y = abs(V1))) + 
  geom_bar(stat = 'identity', aes(fill = tool), position = 'dodge') +
  coord_flip() + 
  labs(y = 'Correlation with Tumor Mutation Burden') + 
  theme(legend.position = 'none', axis.title.y = element_blank(), 
        plot.title = element_text(hjust = 0.5)) 
# immune phenotypes 
dt <- do.call(rbind, lapply(names(LFs_imtrp), function(tool) {
  M <- LFs_imtrp[[tool]]$mariathasan.gex
  # find LFs predictive of immune phenotypes 
  dt <- get_factor_specific_variables(M, colData$`Immune phenotype`)[order(padj)]
  dt$tool <- tool
  return(dt)
}))

dt[,.SD[which.min(padj)],by = c('ref_cl', 'tool')]
dt$tool <- toupper(gsub("_factors", "", dt$tool))
p3 <- ggplot(dt[,.SD[which.min(padj)],by = c('ref_cl', 'tool')], 
             aes(x = ref_cl, y = -log10(padj))) + 
  geom_bar(stat = 'identity', aes(fill = tool), position = 'dodge') +
  geom_hline(yintercept = -log10(0.05)) + 
  labs(x = 'Tumor Immune Phenotypes') + 
  coord_flip() + 
  theme(legend.title = element_blank(), axis.title.y = element_blank(), 
        plot.title = element_text(hjust = 0.5), 
        legend.key.size = unit(0.2, units = 'in'))

# test comparison of maui LFs and published markers from Mariathasan et al
# in terms of response prediction
# published markers from Mariathasan et al (https://www.nature.com/articles/nature25501#Sec4)
# 1. TMB (tumor neoantigen burden) 2. CD8 effector T cell markers (8 genes)
# 3. F-TBRS: fibroblast tgf-beta response signature (4 genes see https://www.nature.com/articles/nature25501/figures/3)

# find top maui LFs informative for response prediction 
M <- scale(LFs_imtrp$maui_factors$mariathasan.gex)
labels <- colData[match(rownames(M), colData$bcr_patient_barcode),]$binaryResponse
df <- cbind(data.frame(M), 'y' = labels)
df <- df[!is.na(df$y),]
df$y <- as.factor(df$y)
# build an initial glmnet model using all LFs on training data
fit <- predict_categorical(df, 'CR/PR') 
# get subset of top LFs 
v <- caret::varImp(fit)[['importance']]
top_lfs <- data.table(v, keep.rownames = T)[order(Overall, decreasing = T)][Overall > 0]$rn

# make a heatmap of top predictive lfs 
m <- as.matrix(df[,-ncol(df)])
f <- ifelse(df$y == 'CR/PR', 'Responders', 'Non-responders')
p4 <- pheatmap::pheatmap(get_basis_matrix(m[,top_lfs], f), silent = T, cellheight = 50, cluster_rows = F)

# survival stratification using the top LF 
require(survminer)
require(survival)
top_lf <- top_lfs[1]
cp <- find_survival_cutpoint(m[,top_lf,drop = F], colData, event_col = 'censOS', duration_col = 'os')
p5 <- plot_survival(surv = colData, 
                    partition.df = data.frame('group' = ifelse(m[,top_lf] > cp, 
                                              paste0(top_lf, ">", round(cp, 1)),
                                              paste0(top_lf, "<=", round(cp, 1))), 
                                              row.names = rownames(m)), 
                    duration_col = 'os', event_col = 'censOS')

# get the assay data for the studied markers for CD8 effectors + TBRS etc.
signatures <- list('cd8_Teff' = human_gene_signatures$`CD 8 T effector`, 
                   'F_TBRS' = c('ACTA2', 'COL4A1', 'TAGLN', 'SH3PXD2A'))
# make a data frame of marker genes + TMB measures + top MAUI LFs
DF <- merge(data.frame(t(gex_all[unlist(signatures),])), 
            data.frame(M[,top_lfs]), by = 'row.names')
DF$TMB <- colData[match(DF$Row.names, colData$bcr_patient_barcode),]$`Neoantigen burden per MB`
# response variable
DF$y <- colData[match(DF$Row.names, colData$bcr_patient_barcode),]$binaryResponse
# remove samples without response variable
DF <- DF[!is.na(DF$y),]
# remove samples without TMB values 
DF <- DF[!is.na(DF$TMB),]
DF <- data.frame(DF[,-1], row.names = DF$Row.names)
DF$y <- as.factor(DF$y)

L <- list('TMB' = 'TMB',
          'MAUI' = top_lfs, 
          'CD8_Teff + F_TBRS' = c(signatures$cd8_Teff, signatures$F_TBRS))
# now predict response using subsets of features 
# get combination of models
combs <- c(asplit(combn(names(L), 2), 2),
           asplit(combn(names(L), 3), 2), 
           'MAUI',
           'CD8_Teff + F_TBRS')
results <- do.call(rbind, lapply(combs, function(x) {
  comb <- paste0(x, collapse = " + ")
  message("trying combination:",comb)
  included <- as.vector(unlist(L[x]))  # variables included in the model
  message("including variables:",paste(included, collapse = ' '))
  res <- predict_cluster_labels_glmnet(scale(as.matrix(DF[,intersect(colnames(DF), included),drop = F])), 
                                DF$y, repeats = 5, sampling = 'down')
  res$comb <- comb
  return(res)
}))

results$comb <- stringr::str_wrap(results$comb, width = 12)
p6 <- ggplot(results[order(auc)][label == 'CR/PR'], 
       aes(x = reorder(comb, -auc), y = auc)) + 
  geom_bar(stat = 'identity', aes(fill = auc)) + 
  geom_errorbar(aes(ymin = auc.lower, ymax = auc.upper), stat = 'identity', 
                width = 0.1) + 
  labs(x = 'Marker Combination', y = 'AUC') + 
  theme(legend.position = 'none', axis.text.x = element_text(angle = 30, hjust = 1))

p5 <- p5$plot + theme(legend.title = element_blank(), legend.direction = 'vertical')
Fig <- cowplot::plot_grid(
  cowplot::plot_grid(p1, p3, p2, nrow = 1, labels = c('A', 'B', 'C')), 
  cowplot::plot_grid(p5, p4$gtable, nrow = 1, labels = c('D', 'E')), 
  cowplot::plot_grid(cowplot::ggdraw(), p6, cowplot::ggdraw(), rel_widths = c(1, 5, 1), nrow = 1, labels = c('', 'F', '')), 
  nrow = 3
)
fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.immunotherapy_response.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(filename = file.path(FigureFolder, fig.name), 
       plot = Fig, units = 'in', 
       width = 12, height = 14)
Legends[[fig_count]] <- "Predicting responders to anti-PD-L1 treatment in a cohort of Metastatic Urothelial
Cancers (N=216) using MAUI. A) Comparison of top latent factors predictive of complete/partial response (CR/PR) to the anti-PD-L1. B) Comparison of top latent factors predictive of immune phenotypes (inflamed/excluded/desert). C)  Comparison of top latent factors with respect to their correlation with tumor mutation burden status of the samples in the cohort D) Survival stratification of the cohort based activation score of top MAUI latent factor (fingerprint) E) Average value of top MAUI latent factors (fingerprints) predictive of response to anti-PD-L1 response groups.
F) Comparison of marker combinations in terms of accuracy in predicting anti-PD-L1 response. TMB: tumor mutation burden, 
CD8_Teff: 12 gene signature for CD8+ effector T-cells, F_TBRS: 4-gene pan-fibroblast TGF-beta response signature, MAUI: top predictive MAUI fingerprints as seen in panel E). Each bar plot represents prediction accuracy (AUC) when using the corresponding subset of variables in an Elastic Net model."
```

```{r tmz_gbm}
# TMZ resistance in GBM anaysis by Jona.
# Here I import figure data to have consistent figure styles
folder <- '/data/local/buyar/arcas/pancancer_multiomics_manuscript/results/tmz_gbm_jona'

# cross-validation results predicting TMZ resistance using 
# maui LFs and other known markers 
aucs <- data.table::fread(file.path(folder, 'aucs.csv'))
aucs$group <- ifelse(grepl('LF', aucs$method), 'maui', 'no-maui')
aucs$method <- gsub("LF[Ss]", "MAUI LFs", aucs$method)
colnames(aucs)[2] <- c('AUC')

p1 <- ggpubr::ggbarplot(aucs, x = "method", y = 'AUC', add = 'mean_se',
                  fill = 'group') +
  coord_flip() + 
  theme(axis.title.y = element_blank(), legend.position = 'none') + 
  scale_fill_brewer(type = 'qual', palette = 3)

# umap responders
umap <- data.table::fread(file.path(folder, 'umap_coords_and_response.csv'))
colnames(umap)[2:3] <- c('UMAP1', 'UMAP2')
#drop samples without umap coords
umap <- umap[!is.na(UMAP1)][!is.na(UMAP2)]
umap$response <- ifelse(umap$response == TRUE, 'responder', 'non-responder')
umap[is.na(response)]$response <- 'unavailable'
umap[is.na(response)]$color <- 'gray'
p2 <- ggscatter(umap, x = 'UMAP1', y = 'UMAP2', color = 'response', size = 2)  + 
  scale_color_manual(values = c('dodgerblue', 'firebrick', 'gray')) + 
  theme(legend.title = element_blank())

# survival plots
gbm_surv <- read.csv(file.path(folder, "survival_time_response_and_predictions.csv"), row.names = 1)
gbm_surv <- gbm_surv[,c(3, 5:7)]
panel_titles <- c('observed response', 'MAUI + MGMT', 
                  'MGMT-expression', 'MGMT-methylation')
plots <- sapply(simplify = F, 1:ncol(gbm_surv), function(i) {
  x <- colnames(gbm_surv)[i]
  plot_survival(surv, partition.df = gbm_surv[,x,drop = F], duration_col = 'PFI.time', event_col = 'PFI')[['plot']] + 
    ggtitle(label = panel_titles[i]) + theme(legend.direction = 'vertical', legend.title = element_blank(), 
                                             legend.position = c(0.5, 0.8))
})

Fig <- cowplot::plot_grid(cowplot::plot_grid(p1, p2, labels = c('A', 'B')),
                          cowplot::plot_grid(plotlist = plots, ncol = 2, labels = c('C', 'D', 'E', 'F')), 
                          nrow = 2, rel_heights = c(1.5, 2))

fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.tmz_gbm.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name),
              plot = Fig, units = 'in', width = 9, height = 11)
Legends[[fig_count]] <- "Predicting TMZ resistance in GBM patients. A) Area under the ROC curve (AUC) for 5-fold cross-validation predictions of response to TMZ. MAUI multi-omic fingerprints outperform both MGMT expression and MGMT promoter methylation, and the best performance is achieved by combining MAUI fingerprints with MGMT expression. B) 2D UMAP projection of TCGA-GBM samples based on MAUI multi-omic fingerprints which are predictive of TMZ response. Responders and non-responders tend to cluster separately. C) Kaplan-Meier curves for progression-free interval for responders and non-responders to TMZ. Response defined as progression-free interval of at least six months, non- or partial-response defined as tumor progression within six months of treatment. D) Kaplan-Meier curves for progression-free intervals for predicted responders and non-responders using MAUI multi-omic fingerprints combined with MGMT expression. E) Kaplan-Meier curves for progression-free intervals for predicted responders and non-responders using MGMT expression. F) Kaplan-Meier curves for progression-free intervals for predicted responders and non-responders using MGMT promoter methylation."

# supplementary figure
topGenes <- data.table::fread(file.path(folder, 'top100_genes.csv'))
colnames(topGenes)[4] <- 'literature'
topGenes$label <- 'Other'
topGenes[MMR == TRUE]$label <- 'MMR Pathway'
topGenes[literature == TRUE]$label <- 'Literature Support'

supFig <- ggpubr::ggbarplot(topGenes, x = 'V1', y = 'Gene score', 
                  fill = 'label') +
  theme(axis.title.y = element_blank(), legend.title = element_blank(),
        axis.text.y = element_text(size = 8)) +
  coord_flip()

supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- supFig
fig.name <- paste0('SuppFigure', supFig_count, '.tmz_gbm.supp.pdf')
names(SuppFigures)[supFig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = supFig, units = 'in', width = 8.27, height = 11.69)
SuppLegends[[supFig_count]] <- "Top 100 genes sorted by importance in predicting TMZ resistance in GBM patients."
```

```{r}
# Show subtypes depend on the objective (dependent variable) - they are dynamic, not universal or carved in stone 
# for a given cancer. There is not just one set of subtypes for a given cancer. Subtypes need to be phrased 
# along with the dependent variable used to come up with the subtypes. 
# Patients clustering depend on the objective of stratification
clin_factors <- c('age_at_initial_pathologic_diagnosis', 'gender', 'ajcc_pathologic_tumor_stage')
project <- 'pan-gastrointestinal'
datatype <- 'cnv_gex_meth_mut'
experiment <- paste0(project, '.', datatype)

M <- LFs$maui_factors[[experiment]]
clin <- process_clinical_covariates(rownames(M))
clin <- subset(clin, select = intersect(colnames(clin), clin_factors))

pfi_vars <- get_prognostic_covariates(surv, clin, M, event_col = 'PFI', duration_col = 'PFI.time')
pfi_clust <- get_kmeans_subtypes(M[,unique(pfi_vars$variable)], k_vals = 3:8, min_cluster_size = 0)
pfi_k <- names(which.min(sapply(colnames(pfi_clust), function(x) plot_survival(surv, pfi_clust[,x,drop = F], 'PFI.time', 'PFI', T))))

os_vars <- get_prognostic_covariates(surv, clin, M, event_col = 'OS', duration_col = 'OS.time')
os_clust <- get_kmeans_subtypes(M[,unique(os_vars$variable)], k_vals = 3:8, min_cluster_size = 0)
os_k <- names(which.min(sapply(colnames(os_clust), function(x) plot_survival(surv, os_clust[,x,drop = F], 
                                                                             'OS.time', 'OS', T))))

# combine with known pan-gi subtype annotation
df <- merge(pfi_clust[,pfi_k,drop = F], os_clust[,os_k,drop = F], by = 'row.names')
df$subtype <- tcga_subtypes[match(df$Row.names, bcr_patient_barcode)]$Subtype_Selected
df[is.na(df$subtype),]$subtype <- 'Unavailable'
df$msi <- MSI[match(df$Row.names, bcr_patient_barcode)]$MSI
df[is.na(df$msi),]$msi <- 'Unavailable'
rownames(df) <- df$Row.names
df$Row.names <- NULL
colnames(df) <- c('Overall Survival', 'Progression Free Survival', 'Molecular Subtype', 'MSI Status')
df <- data.frame(apply(df, 2, as.factor), check.names = F)
df <- df[order(df[,1]),]
m <- apply(df, 2, function(x) as.numeric(as.factor(x)))
rownames(m) <- rownames(df)

colors <- structure(1:7, names = c("1", "2", "3", "4", "5", "6", "7"))
p1 <- Heatmap(t(m), name = "cluster", col = colors, cluster_rows = F, cluster_columns = F,
              show_column_names = F, row_names_side = 'left',
              column_title = 'Pan-gastrointestinal Cancer Samples\n(ordered by Overall Survival Clusters)')
p1 <- grid::grid.grabExpr(ComplexHeatmap::draw(p1))
# compute pairwise AMI 
ami <- sapply(colnames(df), function(x) {
  sapply(colnames(df), function(y) {
    df <- df[!is.na(df[,x]) & !is.na(df[,y]),]
    aricode::AMI(df[,x], df[,y])
  })
})

p2 <- ggcorrplot::ggcorrplot(ami, legend.title = 'AMI', lab = T, type = 'lower', outline.color = 'white') +
  scale_fill_gradient(limits = c(0,1), low = 'white', high = 'red') + 
  guides(fill=guide_legend(title='AMI'))

Fig <- cowplot::plot_grid(p1, p2, labels = 'AUTO', nrow = 2)
supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- Fig
fig.name <- paste0('SuppFigure', supFig_count, '.subtype_dependency.pdf')
names(SuppFigures)[supFig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = Fig, units = 'in', width = 18, height = 12)
SuppLegends[[supFig_count]] <- "Demonstration of how cancer subtypes are context/objective dependent.
A) Heatmap of patient cluster memberships based on different objectives: optimum stratification based on overall survival or progression free survival, molecular homogeneity (TCGA pan-GI subtypes), or micro-satellite instability (MSI) status. Columns are ordered first by membership in the optimum Overall Survival cluster membership. B) Pairwise Adjusted Mutual Information (AMI) of cluster labels obtained via different patient stratification methods. The AMI value ranges from 0 to 1 and the higher the value, the more overlap between patient partitions."
```

```{r print_figures, results='asis', include=TRUE}
base_url <- 'https://bimsbstatic.mdc-berlin.de'
for(i in 1:length(Figures)) {
  fig.url <- gsub("/data/bimsbstatic/public", base_url, file.path(FigureFolder, names(Figures)[i]))
  cat("# Figure",i,"[high-res image](",fig.url,")\n\n")
  print(Figures[[i]])
  cat("\n\n Figure",i,":",Legends[[i]],"\n\n")
}

for(i in 1:length(SuppFigures)) {
  fig.url <- gsub("/data/bimsbstatic/public", base_url, file.path(FigureFolder, names(SuppFigures)[i]))
  cat("# Supplementary Figure",i,"[high-res image](",fig.url,")\n\n")
  print(SuppFigures[[i]])
  cat("\n\n Supplementary Figure",i,":",SuppLegends[[i]],"\n\n")
}
```

`r date()`